---
title: "Week9Assignment"
author: "Bin Lin"
date: "2016-10-30"
output: html_document
---

Assignment Instruction: The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs. You'll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.


1. First step of this assignment is to obtain a API key from NYT website. I am choosing the Books API. This API provides information about book reviews and The New York Times bestsellers lists.

```{r, eval=TRUE}
#I need to install the necessary packages into this markdown file. 
#install.packages("jsonlite")
#install.packages("RCurl")
install.packages('curl', repos = "http://cran.us.r-project.org" )
library(jsonlite)
library(dplyr)
```



2. According to the "README" file on the website, to retrieve best-seller lists, people can use this URL structure.

"http://api.nytimes.com/svc/books/{version}/lists/[date/]{list-name}[.response_format]?[optional-param1=value1]&[...]&api-key={your-API-key}". Curly braces {} indicate required items. Square brackets [] indicat optional items.

README file tells us the version should be v3, response format is .json. However, there is no list-name we can find. Therefore, I was trying to utilize the API for best seller list name to get the type of books that is stored in the Book API.   
I used the function fromJSON from jsonlite to read the content of Book API and converts it to R data frames.
```{r}
url <- "http://api.nytimes.com/svc/books/v3/lists/names.json?api-key=b3b20ef4ff3a4a0ca48764fd9d73e0e5"
json_data <- fromJSON(url)
str(json_data)
```


3. From the code above, we realize the column results contains the most important information about the best selling books. In addition, it tells us the list name for each book is store in the list_name variable. I picked the type of book that I feel interesting, which is Hardcover Business Books.  
```{r}
raw_data <- json_data$results
colnames(raw_data)
raw_data$list_name
```



4. I pass the list name into the API structure again. The following code is doing pretty much the same thing to investigate into what variables that is contained in this API. 
```{r}
url <- "http://api.nytimes.com/svc/books/v3/lists/Hardcover-Business-Books.json?api-key=b3b20ef4ff3a4a0ca48764fd9d73e0e5"
json_data <- fromJSON(url)
raw_data <- json_data$results$books
colnames(raw_data)
```



5. I used the dplyr package to subset the variables that is mostly correlated with best selling books. Especially the ranking for each book. Unfortuantely, NYT only allows me to achieve the top 15 books. Also, I expect there should be some non-zero values for rank_last_week and weeks_on_list, so that I can perform some analysis, but both of the two variables are 0 for all 15 books. It could have been due to the missing data here.
```{r}
books <- raw_data%>%
  select(rank, rank_last_week, weeks_on_list, publisher, title, author)
books
```