---
title: "Final Project"
author: "Bin Lin"
date: "2016-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Introduction:

Since 2014, US has been spending more than $3 trillion dollars annually on healthcare and the average health expenditure is about $9523 per capita. In addition, these numbers keep increasing as those baby boomers start hitting their retirement age. If america spends too much resources on healthcare, there will be less money and resources we can spend on elsewhere because the budget is always limited. 

Medicare and Medicaid are two of the largest federal entitlement programs. People who enroll under Medicare are usually elderly who are over 65 years old. On the other hand, the medicaid is designed for people whose household incomes are under certain limit of federal poverty level. 


2. Objectives
For this final project, I want to create a markdown file that can show the breakdown of spending of medicare and medicaid. In addition to that, I want to compare the spendings between different states, to get some insights why some states have higher spending than the other and investigate if the differences are statistically significant. 



3. Data Sources:
I am using the datasets directly from https://www.data.gov/. The datasets were recently updated, therefore the analysis will be considered current to reflect the utilization of Medicare and Medicaid. Also since I am conducting analysis that is corresponding to the entire country. Govenment data is the best shot for me to obtain.


4. Analysis

a. The first dataset I load to the RStudio is called "Medicare Hospital Spending by Claim". The data  shows average spending levels during hospitals episodes. An MSPB (Medicare Spending per Beneficiary (MSPB) episode includes all Medicare Part A and Part B claims paid during the period from 3 days prior to a hospital admission through 30 days after discharge. The payment amount have been adjusted based on geographic effetcs on payment.  
```{r}
#Loading necessary libraries to RStudio
library(stringr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(jsonlite)
library(XML)
library(RCurl)
library(RMySQL)
library(ggmap)
library(gridExtra)

setwd("C:/Users/blin261/Desktop/DATA607/DATA607Final")

#Exploring the dataset
raw_data <- read.table("Medicare_Hospital_Spending_by_Claim.csv", sep = ",", stringsAsFactors = FALSE, header = TRUE)
head(raw_data)
str(raw_data)

```

This subset of original data frame contains the aggragate information about medicare expenses incurred from hospital visit. I performed neccessary cleaning and transformation of the data. To have it ready for further studies.     
```{r}
raw_complete <- subset(raw_data, raw_data$Period == "Complete Episode")

complete_episode <- raw_complete[, c("State", "Period", "Claim.Type", "Avg.Spending.Per.Episode..State.", "Avg.Spending.Per.Episode..Nation.")]

colnames(complete_episode) <- c("State", "Period", "Claim_Type", "Avg_Spending_Per_Episode_State", "Avg_Spending_Per_Episode_Nation")
head(complete_episode)

```


I grouped the data by state, therefore it is easier to make any comparison between states. I also order the list by descending order according to the average spending per episode in that state.
```{r}
episode_cost <- complete_episode%>%
  group_by(State)%>%
  mutate(Percentage = as.numeric(sub("\\$", "", Avg_Spending_Per_Episode_State)) / as.numeric(sub("\\$", "", Avg_Spending_Per_Episode_Nation)), count = n())%>%
  unique()%>%
  arrange(desc(Percentage))
head(episode_cost)

```



The result of the bar plot imply that states such as Nevada, Texas, Utah, and New Jersey has much higher percentage of medicare hospital spending than the national average. NY which is my home state has actually lower percentage. West Virginia is the state with the lowest medicare hospital spending per member in the entire country. 

```{r}
ggplot(data = episode_cost, aes(x = reorder(episode_cost$State, -episode_cost$Percentage), y = episode_cost$Percentage, fill = State)) + geom_bar(stat = "identity") + coord_cartesian(ylim = c(0.9, 1.1)) + ggtitle("Percentage of Average Spending Per Episode by States")+ xlab("States") + ylab("Percentage to the National Average")
```


Then I pick California and New York to investigate what hospital claims to cause one state (CA) to have higher expenses than the other(NY)
```{r}

#Get a new subset to contain variable about claim type.
claim <- raw_data[, c("Hospital.Name", "State", "Period", "Claim.Type", "Avg.Spending.Per.Episode..State.", "Avg.Spending.Per.Episode..Nation.")]

colnames(claim) <- c("Hosital_Name", "State", "Period", "Claim_Type", "Avg_Spending_Per_Episode_State", "Avg_Spending_Per_Episode_Nation")   

head(claim)

#I also add two new variables to the subset. One is the average cost per claim. It differs based on the claim type. The second variable is the percentage compare to the national average about the medicare spending of each claim type.

claim_cost_NY <- claim%>%
  filter(State == "NY")%>%
  group_by(State, Claim_Type)%>%
  summarize(count = n(), ave_cost = sum(as.numeric(sub("\\$", "", Avg_Spending_Per_Episode_State))) / count, percentage = ave_cost / (sum(as.numeric(sub("\\$", "", Avg_Spending_Per_Episode_Nation)))/ count))
claim_cost_NY

claim_cost_CA <- claim%>%
  filter(State == "CA")%>%
  group_by(State, Claim_Type)%>%
  summarize(count = n(), ave_cost = sum(as.numeric(sub("\\$", "", Avg_Spending_Per_Episode_State))) / count, percentage = ave_cost / (sum(as.numeric(sub("\\$", "", Avg_Spending_Per_Episode_Nation)))/ count))
claim_cost_CA  

```


The following bar plot shows the Medicare program in California has higher spending percentage for almost all claim types except the claims for the nursing facility. While NY and CA's medicare hospital spending are about the same for most claim type, however, california's medicare program spend much more money on the hospice care of the elderly. It outspends NY by almost 30%. 
```{r}
claim_cost <- rbind(claim_cost_NY, claim_cost_CA)
ggplot(data = claim_cost, aes(x = Claim_Type, y = percentage, fill = State)) + geom_bar(stat = "identity", position = "dodge") + ggtitle("NY and CA Claim Cost Comparison")+ xlab("Claim Type") + ylab("Percentage to the National Average")

```



b. Second dataset is about amount of reimbursement that are paid by state Medicaid program for each prescription drugs. It contains variable such as drug name, number of units reimbursed, amount of reimbursement et cetera. 
```{r}
#Load and explore the data. 
setwd("C:/Users/blin261/Desktop/DATA607/DATA607Final")
raw_data1 <- read.table("State_Drug_Utilization_Data_2016.csv", sep = ",", stringsAsFactors = FALSE, header = TRUE)

head(raw_data1)
str(raw_data1)
```

I subset the data and extract the geographic information of each individual state as well as the total drug cost for that state. 
```{r}
state_drug_cost <- raw_data1%>%
  group_by(State, X_latitude, X_longitude)%>%
  summarize(total_drug_cost = sum(Medicaid.Amount.Reimbursed, na.rm = TRUE))%>%
  filter(total_drug_cost != 0 & !is.na(X_latitude) & !is.na(X_longitude))
state_drug_cost
```


Created a visualization using the USA map. It gives us clear picture about prescription expenses from Medicaid program across the country. There are a few states stands out, such as NY, CA, and FL, which are the  three states with the most population in america excluding TX. Just by eye balling this figure. NY's Medicaid seem to be the one with the highest prescription spending.
```{r}
usa <- map_data("state")
ggplot() +
geom_path(data = usa, aes(x = long, y = lat, group = group)) +
geom_point(data = state_drug_cost, aes(x = X_longitude, y = X_latitude, size = total_drug_cost), color = "red")
```


Next few line of codes is just about transformation of the data. I calculated the total product cost of each medication for each states(By the way, state "XX" means the entire country) and the number of that medications were dispense in that state. After we obtain these two numbers we can simply divide the two numbers to calculate the cost of the one unit of that medication. 


For the sake of testing the difference in terms of each medication's cost across states level, Average cost for each unit of medication and its corresponding standard deviation for each states were also calculated.
```{r}
total_cost <- raw_data1%>%
  group_by(State, Product.Name)%>%
  summarize(product_cost = sum(Medicaid.Amount.Reimbursed, na.rm = TRUE), count = sum(Units.Reimbursed, na.rm = TRUE))%>%
  filter(product_cost != 0 & count != 0)%>% 
  arrange(desc(product_cost))

total_cost <- total_cost%>%
  mutate(ave_drug_cost = product_cost / count, average = mean(ave_drug_cost), sd = sd(ave_drug_cost))
head(total_cost, 10)
```


I just want to compare the drug cost for states where most americans live, I perform the test solely for CA, NY, and FL. The barplot shows that medications for HIV and HCV infections are usually most costly in state Medicaid agency's budget. It is reasonable because a lot of these medications have no generic available. In addition to that, they are life-saving medications. Therefore, even the cost is high, people will still have to pay for them. Medications for diabetes and respiratory disorders also have their spot on the highest cost medication list, probably because those medications are common, therefore, many of these prescriptions are filled nationwide. 
```{r}
NY_meds <- subset(total_cost, total_cost$State == "NY")
NY <- ggplot(head(NY_meds, 10), aes(x = reorder(Product.Name, -product_cost), y = product_cost)) + geom_bar(stat = "identity") + ggtitle("NY Highest Cost Medications")+ xlab("Product Name") + ylab("Product Total Cost")

CA_meds <- subset(total_cost, total_cost$State == "CA")
CA <- ggplot(head(CA_meds, 10), aes(x = reorder(Product.Name, -product_cost), y = product_cost)) + geom_bar(stat = "identity") + ggtitle("CA Highest Cost Medications")+ xlab("Product Name") + ylab("Product Total Cost")

FL_meds <- subset(total_cost, total_cost$State == "FL")
FL <- ggplot(head(FL_meds, 10), aes(x = reorder(Product.Name, -product_cost), y = product_cost)) + geom_bar(stat = "identity") + ggtitle("FL Highest Cost Medications")+ xlab("Product Name") + ylab("Product Total Cost")

grid.arrange(NY, CA, FL, nrow=3, ncol=1)
```


Furthermore, I created a data frame that is suitable for conducting two-way ANOVA test. ANOVA test will pretty much tell people if there are statistically significant differences among the mean of the response variable. Explanatory variable in this case is states. 
```{r}
NY_meds <- NY_meds%>%
  mutate(ave_drug_cost = product_cost / count, mean = mean(ave_drug_cost), sd = sd(ave_drug_cost))
head(NY_meds)

CA_meds <- CA_meds%>%
  mutate(ave_drug_cost = product_cost / count, mean = mean(ave_drug_cost), sd = sd(ave_drug_cost))
head(CA_meds)

FL_meds <- FL_meds%>%
  mutate(ave_drug_cost = product_cost / count, mean = mean(ave_drug_cost), sd = sd(ave_drug_cost))
head(FL_meds)

anova_df <- rbind(NY_meds[,c(1, 5)], CA_meds[,c(1, 5)], FL_meds[,c(1, 5)])
head(anova_df)
```


The result of the test shows even though the average medication cost are varid among these three states. The differences are not statistically significant, because the p-value is 0.497 which is above 0.05 (significance level). Also, the confidence interval across 0. 

```{r}
drug_aov <- aov(anova_df$ave_drug_cost ~ anova_df$State)
summary(drug_aov)
TukeyHSD(drug_aov)
confint(drug_aov)
```


5. Conclusion.
In general, the expenses of hospital episode by Medicare differs quite a lot. Some states are around 9% above the national average, while the other state could be about 7% below national average. By breaking down the expense by different claim type, we can usually detect where the discrepancies are. For example, from the NY and CA claim cost comparison, we know it is the hospice care that account for most of the differences between the two states. For the sake of time, I could not perform the similar analysis among other states.
Another point that arise after analyzing the drug cost data from Medicaid program is that generally speaking, if a state has high proportion of HIV, HCV patients, the state medicaid program will have to reimburse more for the corresponding prescription. This is phenomenon is manifested in the national aggreagate data also the state specific data. Moreover, if we want to compare the prescription cost among different states, It is very difficult to establish statistically significant conclusion, even though the sample data seem to show differences in terms of average drug cost. With over 1 million observations, we still can not claim prescriptions sold in NY is cheaper that sold in FL, although the NY does have sample mean that is about 24 dollars cheaper. More investigation will be undergoing to gain more insight on this issue. 







