---
title: "Project 2"
author: "Bin Lin"
date: "2016-10-9"
output: html_document
---

Project Instruction: Choose any three of the “wide” datasets identified in the Week 6 Discussion items. 
The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work.



Dataset 1: Religion and Income Distribution
Contributor: Yifei Li
Source: Introduction to R. (2013). Retrieved from https://ramnathv.github.io


```{r}
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("ggplot2")
library(dplyr)
library(tidyr)
library(ggplot2)
```


Load the CSV file and transform the data from "wide" to "long"
```{r}
religion_income <- read.csv("C:/Users/blin261/Desktop/DATA607/Religion_Income.csv", header = TRUE, stringsAsFactors = FALSE, check.names=FALSE)
religion_income

long_data <- religion_income%>%
  gather(income_group, frequency, 2:10)
head(long_data)
```


Tidy the data. Get total frequency of income for each individual religion group. I also calculated the percentage of each income group within its religion group.
```{r}
r_i <- long_data%>%
  group_by(religion)%>%
  mutate(total = sum(frequency), percentage = frequency/total)%>%
  arrange(religion)
head(r_i)
```


The graph has shown for each religion group, the income distribution normally peaks at $50k-70k, with smallest porportion of people making lower than $10k. As the income keeps going up above $50k-70k, the proportion of people usaully goes down. This makes sense, in real life, we do not see that much people making over 150k.  
```{r}
r_i$income_group<-ordered(r_i$income_group,levels=c("<$10k","$10-20k","$20-30k","$30-40k","$40-50k","$50-75k","$75-100k", "$100-150k", "$>150k"))

ggplot(data = r_i, aes(x = income_group, y = percentage)) + geom_bar(stat="identity") + facet_wrap(~religion) 
```




Dataset 2: Gaming Jobs and Broadband
Contributor: Bruce Hao
Source: http://www.pewinternet.org/datasets/june-10-july-12-2015-gaming-jobs-and-broadband/


Load the csv file, and subsetting the variables that will help with our analysis. 
```{r}
gaming_job_broadband <- read.csv("C:/Users/blin261/Desktop/DATA607/GamingJobsandBroadband.csv", header = TRUE, stringsAsFactors = FALSE, check.names=FALSE)

gaming <- gaming_job_broadband %>%
  select(game4, emplnw, stud, age, educ2, inc)
head(gaming)
```


The original data contains observations that are mostly numbers, which stand for certain responses. The following code just making those responses more meaningful by changing the data type from numeric to string which is more human readable. Moreover, it is very helpful to order them in a more sensible sequence which will be easier to perform some analysis later on. 
```{r}
gaming$game4[gaming$game4 == 1] <- "gamer"
gaming$game4[gaming$game4 == 2] <- "not_gamer"

gaming$emplnw[gaming$emplnw == 1] <- "full_time"
gaming$emplnw[gaming$emplnw == 2] <- "part_time"
gaming$emplnw[gaming$emplnw == 3] <- "retired"
gaming$emplnw[gaming$emplnw == 4] <- "not_employed"
gaming$emplnw <- ordered(gaming$emplnw, levels = c("full_time", "part_time", "retired", "not_employed"))


gaming$stud[gaming$stud == 1] <- "full_time_student"
gaming$stud[gaming$stud == 2] <- "part_time_student"
gaming$stud[gaming$stud == 3] <- "no"
gaming$stud <- ordered(gaming$stud, levels = c("full_time_student", "part_time_student", "no"))


gaming$educ2[gaming$educ2 == 1] <- "less_than_HS"
gaming$educ2[gaming$educ2 == 2] <- "HS_incomplete"
gaming$educ2[gaming$educ2 == 3] <- "HS"
gaming$educ2[gaming$educ2 == 4] <- "some_college"
gaming$educ2[gaming$educ2 == 5] <- "associate"
gaming$educ2[gaming$educ2 == 6] <- "bachelor"
gaming$educ2[gaming$educ2 == 7] <- "some_postgraduate"
gaming$educ2[gaming$educ2 == 8] <- "post_graduate"
gaming$educ2 <- ordered(gaming$educ2, levels=c("less_than_HS", "HS_incomplete", "HS", "some_college", "associate", "bachelor", "some_postgraduate", "post_graduate"))


gaming$inc[gaming$inc == 1] <- "<$10k"
gaming$inc[gaming$inc == 2] <- "$10k-20k"
gaming$inc[gaming$inc == 3] <- "$20-30k"
gaming$inc[gaming$inc == 4] <- "$30-40k"
gaming$inc[gaming$inc == 5] <- "$40k-50k"
gaming$inc[gaming$inc == 6] <- "$50k-75k"
gaming$inc[gaming$inc == 7] <- "$75k-100k"
gaming$inc[gaming$inc == 8] <- "$100k-150k"
gaming$inc[gaming$inc == 9] <- "$>150k"
gaming$inc <- ordered(gaming$inc, levels = c("<$10k", "$10k-20k", "$20k-30k", "$30k-40k", "$40k-50k", "$50k-75k", "$75k-100k", "$100k-150k", "$>150k"))

head(gaming)
```


Still, the data contains observations that does not belong to our interests. We can use functions in dplyr and tidyr to filter out  any missing values or response that does not help our analysis.
```{r}
gaming <- gaming%>%
  filter(game4 == "gamer" | game4 == "not_gamer")%>%
  filter(emplnw == "full_time" | emplnw == "part_time" | emplnw ==  "retired" | emplnw == "not_employed")%>%
  filter(stud == "full_time_student" | stud == "part_time_student" | stud == "no")%>%
  filter(educ2 == "less_than_HS" | educ2 == "HS_incomplete" | educ2 == "HS" | educ2 == "some_college" | educ2 == "associate" | educ2 == "bachelor" | educ2 == "some_postgraduate" | educ2 == "post_graduate")%>%
  filter(inc == "<$10k" | inc == "$10k-20k" | inc == "$20k-30k" | inc == "$30k-40k" | inc == "$40k-50k" | inc == "$50k-75k" | inc == "$75k-100k" | inc == "$100k-150k" | inc == "$>150k")%>%
  arrange(game4, emplnw)
head(gaming)
View(gaming)
```


The first graph I created compares the relationship between gaming and employment status. We can tell non-gamer has slightly higher percentage of people working full time, and lower percentage of people unemployeed. We also notice there are more people who retired in the not-gamer group. This may be explained by the reason elderly persons may not have quite exposure to internet, computers or smartphones as the young people, therefore, they tend to not playing games.  
```{r}
game_emp <- gaming%>%
  group_by(game4, emplnw)%>%
  summarize(count = n())%>%
  mutate(total = sum(count), percentage = count/total)%>%
  arrange(game4,emplnw)
head(game_emp)
ggplot(data = game_emp, aes(x = emplnw, y = percentage, fill = game4)) + geom_bar(stat="identity") + facet_wrap(~game4)
```


The following graph shows non-gamer makes more money than gamers, as higher proportion of them belong to the higher income group. We can connect this result to the result we got from the first graph. For non-gamers who tend to have full time jobs, of course their income is going to be relatively higher. 
```{r}
game_inc <- gaming%>%
  group_by(game4, inc)%>%
  summarize(count = n())%>%
  mutate(total = sum(count), percentage = count/total)%>%
  arrange(game4,inc)
head(game_inc)
ggplot(data = game_inc, aes(x = inc, y = percentage, fill = game4)) + geom_bar(stat="identity", position = "dodge")
```


This graph just shows non-gamers have relatively higher education level (obtain a degree higher than high school diploma). 
```{r}
game_edu <- gaming%>%
  group_by(game4, educ2)%>%
  summarize(count = n())%>%
  mutate(total = sum(count), percentage = count/total)%>%
  arrange(game4,educ2)
head(game_edu)
ggplot(data = game_edu, aes(x = educ2, y = percentage, fill = game4)) + geom_bar(stat="identity", position = "dodge")
```




Dataset 3: Lending Club Loan Stat 2016Q2
Contributor: Bin Lin
Source: https://www.lendingclub.com/info/download-data.action

The first step is to load the data, apparently from the dimention function, we know it is a very large datasets.
```{r}
lending_club <- read.csv("C:/Users/blin261/Desktop/DATA607/LoanStats_2016Q2.csv", header = TRUE, stringsAsFactors = FALSE)
dim(lending_club)
```


Then I tidy, subset, and transform the data. In the meantime, I created a new variable called loantoincome_ratio, which I think is very important variable for us to gain insight about the loan data.
```{r}
loan_stat <- lending_club %>%
  select(term, grade, loan_amnt, annual_inc, int_rate)%>%
  na.omit()
head(loan_stat)
```


The first graph shows there are way more 36-month loans approved than the 60-month loans. The distribution are both skewed to the right. The most 36-month loans receive B grade while most 60-month loans receive C grade. 
```{r}
loan <- loan_stat %>%
  group_by(term, grade)%>%
  summarize(count = n())
head(loan)

ggplot(data = loan, aes(x = grade, y = count, fill = term)) + geom_bar(stat="identity") + facet_wrap(~term)
```

The second graph tell us most of the loans have loan-to-income ratio less than 50%, probabaly because lending club thoughts this type of loan has lower risk. so that the company will be willing to lend the money to these clients. Another thing we found out is on the 60-month loan group, there are more loans with high interest rate (greater than27.34%) and fewer loans with low interest rate (less than 8.59%)
```{r}
loan <- loan_stat %>%
  mutate(loantoincome_ratio = (loan_amnt)/(annual_inc))
head(loan)

ggplot(data = loan, aes(x = loantoincome_ratio, y = int_rate, color = grade)) + geom_point(stat="identity") + facet_wrap(~term) + xlim(0, 1)
```
