---
title: "Week5Assignment"
author: "Bin Lin"
date: "2016-10-2"
output: html_document
---

(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You're encouraged to use a "wide" structure similar to how the information appears above, so that you can practice tidying and transformations as described below.
(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.
(3) Perform analysis to compare the arrival delays for the two airlines.

```{r}
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("ggplot2")

library(dplyr)
library(tidyr)
library(ggplot2)
```


After creating the csv file in my local machine, I loaded the data file into Rstudio. 
```{r}
flights_delayed <- read.csv("C:/Users/blin261/Desktop/DATA607/flights_delayed.csv", header = TRUE, stringsAsFactors = FALSE)
flights_delayed
str(flights_delayed)
```


From the raw databases, we can tell the dataset is not very tidy and appears to be "wide". Some of the column names do not really make sense. The following code will first change the column names and the observation's name as well.      
```{r}
colnames(flights_delayed)[1:2] <- c("Airlines", "Punctuality")
flights_delayed$Punctuality[flights_delayed$Punctuality == "on time"] <- "on_time"
flights_delayed
```


Then what I am doing with the following code is to transform the dataset from being "wide" to be "long", so that all the data that we are interested in can be shown in a data frame. 
```{r}
flights <- flights_delayed%>%
gather(Destinations, Frequency, Los.Angeles:Seattle)%>%
  spread(Punctuality, Frequency)
flights
```


In order to perform analysis, more variables need to be created, such as the total flights for each airline company on each destinations. In the meantime, we can also calculate the percentage of flights that were delayed. This can be accomplished using function called "mutate". 
```{r}
flights <- flights%>%
  mutate(total = delayed + on_time,Percentage = delayed / total)

flights
str(flights)
```


After we create the new data frame called flights, we want to calculate the aggregate percentage of flights that were delayed for each airline. group_by function can help us calculate variables by group, in this case, the airline companies. summarize function can help us perform calculations in an aggregate scale. The result generated is a table. As shown in the following, AM WEST seems to be a better airline company overall as it has lower flights delayed 
rate compared to ALASKA. 
```{r}
f <- flights%>%
  group_by(Airlines)%>%
  summarize(total_delayed = sum(delayed), total_on_time = sum(on_time), total_flights = sum(total), total_percentage = total_delayed / total_flights)
f
```


We also want to know what is each airline companies' flight delayed rate look like for each destination. Of course, we need to group the data by airline companies again, then use spread function to display each destinations in columnns. The observations of those variables will be percentage of flight delays. Surprisingly, for every destination, it is always the ALASKA which has relatively lower flight delayed rate. 
```{r}
F <- flights%>%
  group_by(Airlines)%>%
  arrange(Destinations, Airlines)%>%
  select(Airlines, Destinations, Percentage)%>%
  spread(Destinations, Percentage)
F
```


If ALASKA is able to beat AW WEST on each destinations in terms of lower flight delayed rates, how can AW WEST still be a better airline company overall? That is a question we want to ask. Actually this phenomenon has a very technical term called Simpson's Paradox. If we want to investigate the reason behind this paradox, it is good idea to draw some graphs to have visual information about these numbers.According to the graph, we are able to know that both companies have very closed flight delayed rates. The majoy discrepancies between the two are the numbers of flights that land at Phoenix. Since most of AW WEST's flights go to Phoenix, in addition, the flights delayed rates have been the lowest among all destinations for both airline companies in Phoenix, AW WESt will be able to take advantage of that to lower down its overall flight delayed rate.   

```{r}
ggplot(data = flights, aes(y = total, x = Destinations))+ geom_point(aes(color = Airlines)) + labs(title = "Total Flights")

ggplot(data = flights, aes(y = Percentage, x = Destinations))+ geom_point(aes(color = Airlines)) + labs(title = "Percentage of Delayed Flights")
```

