---
title: "Bin Lin - Project 3"
author: "Bin Lin"
date: "2016-10-22"
output: html_document
---

```{r}
#install.packages("RMySQL")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("ggplot2")
#install.packages("DBI")

library(DBI)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(RMySQL)




```



Jose's Code
```{r}
JA_Data <- read.csv("https://raw.githubusercontent.com/juddanderman/cuny-data-607/master/Project3/linkedin-profiles-skills.csv", encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = FALSE)
JA_Data <- cbind("LinkedIn", JA_Data[ , c(10,3,4,2,5,6)], NA)
JA_Data[ , 2] <- tolower(JA_Data[ , 2])
JA_Data[ , 2] <- iconv(JA_Data[ , 2], from = "latin1", to = "UTF-8")
JA_Data <- unique(JA_Data)
JA_Data$ID <- seq.int(nrow(JA_Data))
colnames(JA_Data) <- c("Source","Skill","Title","Location","Name","School","Degree","Company","Record_ID")
t(head(JA_Data, 1))

KC_Data <- read.csv("https://raw.githubusercontent.com/cunyauthor/Project3/master/API_Job.csv", encoding="UTF-8", na.strings=c("","NA"), stringsAsFactors = FALSE)
KC_Data <- KC_Data[KC_Data[ , 1] != "count",] # Remove heading rows
KC_Data <- KC_Data[!is.na(KC_Data[ , 5]),] # Remove rows with blank skills
KC_Data <- cbind(Source = "KDnuggets+Dice", KC_Data[ , c(5,7,9)], NA, NA, NA, KC_Data[ , 8])
KC_Data[ , 2]  <- as.character(str_extract_all(KC_Data[ , 2] , "l\\=\\S+\\&c"))
KC_Data[ , 2]  <- str_replace_all(KC_Data[ , 2] , "(l\\=|\\&c)", "")
KC_Data[ , 2]  <- str_replace_all(KC_Data[ , 2] , "\\+", " ")
KC_Data$ID <- seq.int(nrow(KC_Data))
colnames(KC_Data) <- c("Source","Skill","Title","Location","Name","School","Degree","Company","Record_ID")
t(head(KC_Data, 1))

# I just changed the connection to the MySQL, so that the password and user name #won't be shown. 
rmysql.settingsfile<-"C:/ProgramData/MySQL/MySQL Server 5.7/my.ini"
connection <- dbConnect(RMySQL::MySQL(), default.file=rmysql.settingsfile, dbname = "assignment2", user=NULL, password=NULL)

dbSendQuery(connection, 'CREATE SCHEMA IF NOT EXISTS Skills;')
dbSendQuery(connection, 'USE Skills;')
dbSendQuery(connection, 'DROP TABLE IF EXISTS tbl_LinkedIn;')
dbSendQuery(connection, 'DROP TABLE IF EXISTS tbl_KDnuggets_Dice;')

dbWriteTable(connection, "tbl_LinkedIn", JA_Data, append = TRUE, row.names = FALSE)
dbSendQuery(connection, "ALTER TABLE tbl_LinkedIn
            MODIFY COLUMN Record_id MEDIUMINT NOT NULL,
            MODIFY COLUMN Source VARCHAR(25) NOT NULL,
            MODIFY COLUMN Skill VARCHAR(50) NOT NULL,
            MODIFY COLUMN Title VARCHAR(250) NULL,
            MODIFY COLUMN Location VARCHAR(50) NULL,
            MODIFY COLUMN Name VARCHAR(50) NULL,
            MODIFY COLUMN School VARCHAR(75) NULL,
            MODIFY COLUMN Degree VARCHAR(100) NULL,
            MODIFY COLUMN Company VARCHAR(50) NULL,
            ADD PRIMARY KEY (Record_id);")

dbWriteTable(connection, "tbl_KDnuggets_Dice", KC_Data, append = TRUE, row.names = FALSE)
dbSendQuery(connection, "ALTER TABLE tbl_KDnuggets_Dice
            MODIFY COLUMN Record_id MEDIUMINT NOT NULL,
            MODIFY COLUMN Source VARCHAR(25) NOT NULL,
            MODIFY COLUMN Skill VARCHAR(50) NOT NULL,
            MODIFY COLUMN Title VARCHAR(250) NULL,
            MODIFY COLUMN Location VARCHAR(50) NULL,
            MODIFY COLUMN Name VARCHAR(50) NULL,
            MODIFY COLUMN School VARCHAR(75) NULL,
            MODIFY COLUMN Degree VARCHAR(100) NULL,
            MODIFY COLUMN Company VARCHAR(50) NULL,
            ADD PRIMARY KEY (Record_id);")

All_Data <- dbGetQuery(connection, "SELECT * FROM tbl_LinkedIn
                       UNION SELECT * FROM tbl_KDnuggets_Dice
                       ORDER BY Source, Skill, Title;")

```




My Code: 
I created a barplot for the entire dataset, without regarding to linkedin or KD Nuggets or Dice. The graph shows the top 10 skills and their frequency.
```{r}

raw_data <- All_Data[!(is.na(All_Data$Skill) & All_Data$Skill != "character(0)"), ]
a <- which(with(raw_data, table(raw_data$Skill)) > 100)
head(a)


barplot(a[3:12], las=2)

```




I am interested in the rlationship between each company and their most desired employee skills. 
```{r}
raw_data <- All_Data%>%
  select(Company, Skill)%>%
  na.omit()
head(raw_data)


b <-raw_data%>%
  group_by(Company, Skill)%>%
  summarize(count = n())%>%
  mutate(total = sum(count), percentage = count/total)%>%
  filter(total > 100 & percentage > 0.03)%>%
  arrange(desc(total), desc(percentage))
head(b)
```



The first barplot shows different companies actually have different interest of skill sets they want candidates to have. I pick whatever companies have more than 100 positions open. However, I think the better way to do is to get the top 10 companies that are hiring. The percentage I set up is just for the purpose of getting rid off the skills that do not meet most companies' interest.  The second plot looks a little messy. If someone can help me organize it, (for each barplot, the skills can line up from the most frequent to the lleast frequent), I really appreaciate. If not, maybe just take off the second barplot.


I also want to connect skills with each schools. But it is actually a bad idea, because a lot of data is missing scholl information. 

```{r}
ggplot(data = b, aes(x = Skill, y = count, fill = Company)) + geom_bar(stat="identity") + theme(axis.text.x = element_text(angle=90, hjust = 1))


ggplot(data = b, aes(x = Skill, y = count, fill = Company)) + geom_bar(stat ="identity", position = "dodge") + facet_wrap(~Company) +
theme(axis.text.x = element_text(angle = 90, hjust = 1))



```



Jose's Code
```{r}
dbSendQuery(connection, 'DROP TABLE tbl_LinkedIn;')
#dbSendQuery(connection, 'DROP TABLE tbl_Friends;')
dbSendQuery(connection, 'DROP SCHEMA Skills;')
dbDisconnect(connection)
```



